---
title             : "Fundamental errors of data collection & validation undermine claims of 'Ideological Intensification' made by the National Association of Scholars"
shorttitle        : "Flawed data validation by the NAS"

author: 
  - name          : "Emilio M. Bruna"
    affiliation   : "1,2"
#    address       : "Department of Wildlife Ecology and Conservation, University of Florida, PO Box 110430, Gainesville, FL 32611-0430, USA"
    email         : "embruna@ufl.edu"
    # role:
    #   - Methodology
    #   - Data curation
    #   - Investigation
    #   - Funding acquisition
    #   - Conceptualization
    #   - Formal analysis
    #   - Methodology
    #   - Project administration
    #   - Resources
    #   - Software
    #   - Supervision
    #   - Validation
    #   - Visualization
    #   - Writing of the original draft

affiliation:
  - id            : "1"
    institution   : "Department of Wildlife Ecology and Conservation, University of Florida, PO Box 110430, Gainesville, FL 32611-0430, USA"
  - id            : "2"
    institution   : "Center for Latin American Studies, University of Florida, PO Box 115530, Gainesville, FL 32611-5530, USA"

authornote:  |

  Address for correspondence: University of Florida, WEC, PO Box 110430, Gainesville, FL 32611-0430, USA (embruna@ufl.edu)

  The code and data used in this analysis are available for download and improvement at https://github.com/embruna/quantdei_nas. 

# 
# abstract: |
#   Abstract of the paper.
  
  
  
keywords          : "Up,to,eight,keywords"
wordcount         : "X"

bibliography      : "NAS_response.bib"

floatsintext      : no
figurelist        : yes
tablelist         : no
footnotelist      : no
numbersections    : no
linenumbers       : yes
mask              : no
draft             : no
# replace_ampersands: yes

classoption       : "man, donotrepeattitle" # suppresses title on 1st page of MS
output            : papaja::apa6_pdf
mainfont: Times New Roman
fontsize: 12pt
csl: "science.csl"
fig_caption: yes
keep_tex: yes

header-includes:
  - \raggedbottom
  - \usepackage{endfloat} #[nomarkers] excludes the {insert figure x around here] from main text. The others exclude the list of tables and figures. https://cs.brown.edu/about/system/managed/latex/doc/endfloat.pdf
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{tabu}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}

---


```{r setup, include = FALSE}
library("papaja")
r_refs("NAS_response.bib")
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(egg)
library(magick)
```


```{r GlobalOptions, include=FALSE}
options(knitr.duplicate.label = 'allow')
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
# knitr::opts_chunk$set(fig.pos = 'h')
```


```{r cached=TRUE, echo = FALSE, warning=FALSE,message = FALSE}
knitr::opts_chunk$set(echo = FALSE)

tweet_fail<-read_csv("./error_analysis_bruna/validation_output/twitter_notdei.csv",col_types = cols())

twitter_fail_summary<-read_csv("./error_analysis_bruna/validation_output/twitter_notdei_summary.csv",col_types = cols())

fail_tweets_sum<-twitter_fail_summary %>% summarize(n=sum(fail_tweets)) %>% mutate(n=round(n,-3))

grants_nsf_diversity<-read_csv("./error_analysis_bruna/validation_output/grants_nsf_diversity.csv")

fail_tweets_sum<-twitter_fail_summary %>% summarize(n=sum(fail_tweets)) %>% mutate(n=round(n,-3)) 
nsf_div_sum<-grants_nsf_diversity %>% summarize(n=nrow(grants_nsf_diversity)) %>% mutate(n=round(n,-3))
```

# Introduction 



>> \@arizonaalumni: Good luck to former #UofA student and \@nascar champ \@kurtbusch as he attempts to race in both the Indy 500 and Coke 600. #beardown!

Efforts to advance Diversity, Equity, and Inclusion (hereafter, DEI) at universities in the United States have emerged as another contentious issue in an increasingly polarized political climate [@diepFlaGovernorAsked2023; @keldermanPlanDismantle2023]. While individuals and organizations critical of DEI often claim that these programs have become increasingly pervasive and ideological [@goad2022ideological], this assertion is rarely supported with empirical evidence.  
  The National Association of Scholars (i.e., NAS) recently published a report by Mason Goad and Bruce R. Chartwell [@goad2022ideological] which the authors claim is "the largest quantitative study of the growth of DEI-related language in the sciences" published to date. Goad and Chartwell searched university web pages and Twitter accounts, funding agency databases, and repositories for scientific literature to quantify changes over time in the use of "DEI-related terminology" (e.g., "advocacy", "ally", "diversity", "equity", "justice", "privilege", "race"). They claim to show a dramatic increase in the use of these terms in university communications and the scientific literature since 2010, which they view as unambiguous empirical evidence of "ideological intensification" in the academic and scientific arenas [@goad2022ideological]. Goad and Chartwell conclude that if the trends they document continue, "the future of STEM, along with the rest of the academy, is almost certainly imperiled" (see @goad2022ideological, p. 47), so they encourage others to use their data-mining tools and database in their own research. Since the report's release in December 2022, it has been widely hailed and distributed by prominent DEI critics such as Jordan Peterson, Colin Wright, and Christopher Rufo [@petersonJordanbpetersonRetweetNASorg2022;@rufoRealchrisrufoRetweetGoadMason2023; @goadIdeologicalIntensificationSTEM2023].  
  Goad and Chartwell made the laudable decision to post the code used to harvest and process the records used in their report [@nationalassociationofscholarsNationalAssociationScholars2022], along with the "clean" data used to graph the trends on which they base their conclusions [@nationalassociationofscholarsDataQuantitativeStudy2022], "so that other analysts can scrutinize the methods and replicate them" (see @goad2022ideological, p. 48). Having done so [@brunaCodeIdentifyingErrors2023, _S1 Methods_], I conclude that they failed to conduct even the most rudimentary data validation procedures, and that as a result the "clean" data sets with which they conducted their analyses contain thousands of irrelevant records and duplications. Notable examples include the tweet that opened this letter - one of over `r fail_tweets_sum` about topics ranging from sporting events ("race") to members of the Supreme Court ("justice") to hedge funds ("equity", _Tables S1,S2_) - along with almost `r nsf_div_sum` NSF grants for ecological and evolutionary research on species "diversity" (_Table S3_), and hundreds of "DEI articles in STEM journals" that were actually published in outlets such as _Critical Sociology_, _The Medical Law Review_, and _The Annual Review of Law and Social Science_ (_Table S4_).  
  Policy Institutes play a vital role in shaping higher education and research policy, but only when they put scholarship ahead of ideology. Goad and Chartwell's conclusions regarding the prevalence of DEI in STEM were based entirely on visualizations of fundamentally flawed datasets. The NAS should retract their report immediately; failure to do undermines both the integrity and intellectual rigor that they and the report's authors claim to espouse as fundamental principles.


\newpage
# References

::: {#refs custom-style="Bibliography"}
:::




<!-- ```{r, child='Bruna_Supplement.Rmd'} -->
<!-- ``` -->


<!-- \newpage -->

<!-- # (APPENDIX) Appendix {-} -->

<!-- ```{r child = "Bruna_Supplement_Tables.Rmd"} -->
<!-- ``` -->

<!-- ```{r child = "./error_analysis_bruna/response_ms/Bruna_Supplement.Rmd"} -->
<!-- ``` -->


